Demography and selection since maize domestication

### 2015-05-05 manuscript.tex
updates

\dropcap{T}his is the beginning of the article. Notice the
dropcap... that is a neat feature that PNAS likes, I think it looks
pretty neat too! This is where the introduction will live.

\section{Results}
\subsection{Patterns of variability differ between genic and
  nongenic regions of the genome}
Previous research of the maize domestication process has relied upon
observations drawn from genic DNA \textcolor{red}{(several references
  here)}. Our data were generated through whole genome sequencing,
which eliminated this constraint. Importantly, we observed substantial
differences in patterns of diversity between genic and non-genic regions
of the genome for both maize and teosinte. For maize, mean pairwise
diversity ($\pi$) within genes was significantly lower than at
positions at least 5kb away from genes (0.00668 within, 0.00691 away, p$<$2e-44). The same
pattern of significantly greater diversity outside of genes was observed in
teosinte, but to a larger extent. For teosinte we observed $\pi$
within genes to be 0.0088 and at least 5kb from genes it was 0.1153
(p$\approx$0). These obserrvations suggest that genes are not evolving
neutrally in maize or teosinte. Instead, some form of selection must
be reducing diversity within genes. Additionally, these observations
suggest that demographic inference, which relies on the assumption of
neutrally evolving DNA, will be more accurate if it is based on
observations from non-genic genetic material.

\subsection{Hard sweeps do not shape maize (or teosinte) diversity}
A mutation that is immediately beneficial and positively selected leaves a classical hard
sweep signature in the genome, whereby genetic diversity surrounding
that mutation is reduced as the haplotype where the mutation first
arose increases in frequency to fixation. The prevalence of such hard
sweeps was evaluated by comparing diversity
surrounding non-synonymous and synonymous substitutions in maize and
teosinte since divergence from tripsicum, roughly \textcolor{red}{ZZZ} years before
present \textcolor{red}{(ref here)}. For both taxa, no difference in diversity
surrounding these classes of substitutions was identified (Figure \ref{hardSweeps}). Additionally, diversity around maize substitutions
not seen in teosinte was investigated. These sites have the
potential to correspond to recent maize sweeps, occuring after the
split from teosinte. Again, no difference was observed
between diversity around synonymous and nonsynonymous
substitutions \textcolor{red}{(Figure here)}. Together, these observations suggest that hard sweeps
are not a primary form of selection for either maize or teosinte.

@@

\subsection{Patterns of purifying and background selection can be explained by demography}
Purifying selection refers to the situation where deleterious
mutations arising in a population are continuously selected against. When this
form of selection is operating, it can serve to reduce genetic
variability at linked neutral sites, a phenomenon referred called
background selection \cite{charlesworth1993}. Purifying and background selection lead
to lower diversity within genes and other functional sites relative
to neutral regions \textcolor{red}{(ref here)}. We investigated purifying
selection in maize and teosinte by evaluating the average magnitude of reduced
diversity within genes and recovery away from genes in both taxa
\textcolor{red}{(figure here)}. When standardized by neutral
levels (diversity far from genes), a stronger reduction of diversity
and slower recovery was observed for teosinte than for maize, implying
that purifying selection has left a more pronounced signature in the
teosinte genome. This conflicted with our \emph{a priori} hypothesis;
we expected that strong artificial selection since domestication would
have caused enhanced purifying selection for maize. We therefore
conducted the same analysis based on singleton diversity. As a class,
singleton alleles depict the most recent patterns of evolution, but
also have the lowest effect on pairwise diversity. Therefore, unlike
pairwise diversity, patterns of singleton
diversity reflect recent patterns of evolution. When
evaluating the data in this manner, an opposite relationship was
observed. Maize singleton diversity was lower than teosinte singleton
diversity near genes, and recovered more slowly
\textcolor{red}{(figure here)}, implying that in the
recent past maize has been more influenced by purifying selection than
teosinte. Together, these findings imply that demographic history has
a strong influence on the effect of purifying selection. Historically,
teosinte has had a larger population size than maize, and only
recently has maize population size overcome that of teosinte. Since
the efficacy of purifying selection scales with population size,
these results likely reflect changes in $N_e$ more than they reflect
underlying changes in selection pressure.

@@

\bibliography{Reference}
\bibliographystyle{abbrvnat}
%\nocite{*}

>>>>

\dropcap{T}his is the beginning of the article. For now, I will give a
bulleted list of the introduction:
\begin{enumerate}
\item Maize
\item Plant domestication
\item Maize domestication
\item Importance of demography
\item Background selection
\end{enumerate}

\section{Results}
\subsection{Patterns of variability differ between genic and
  nongenic regions of the genome}
Previous research of the maize domestication process has relied upon
observations drawn from genic DNA \textcolor{red}{(several references
  here)}. Our data were generated through whole genome sequencing,
which eliminated this constraint. Importantly, we observed substantial
differences in patterns of diversity between genic and non-genic regions
of the genome for both maize and teosinte. For maize, mean pairwise
diversity ($\pi$) within genes was significantly lower than at
positions at least 5kb away from genes (0.00668 within, 0.00691 away, p$<$2e-44). The same
pattern of significantly greater diversity outside of genes was observed in
teosinte, but to a larger extent. For teosinte we observed $\pi$
within genes to be 0.0088 and at least 5kb from genes it was 0.1153
(p$\approx$0). These obserrvations suggest that genes are not evolving
neutrally in maize or teosinte. Instead, some form of selection is likely
reducing diversity within genes. Additionally, these observations
suggest that demographic inference, which relies on the assumption of
neutrally evolving DNA, will be more accurate if it is based on
observations from non-genic genetic material.

\subsection{Hard sweeps do not shape maize (or teosinte) diversity}
A mutation that is immediately beneficial and positively selected leaves a classical hard
sweep signature in the genome, whereby genetic diversity surrounding
that mutation is reduced as the haplotype where the mutation first
arose increases in frequency to fixation. The prevalence of such hard
sweeps was evaluated by comparing diversity
surrounding non-synonymous and synonymous substitutions in maize and
teosinte since divergence from tripsicum, roughly \textcolor{red}{ZZZ} years before
present \textcolor{red}{(ref here)}. For both taxa, no difference in diversity
surrounding these classes of substitutions was identified (Figure \ref{hardSweeps}). Additionally, diversity around maize substitutions
not seen in teosinte was investigated. These sites have the
potential to correspond to recent maize sweeps, occuring after the
split from teosinte. Again, no difference was observed
between diversity around synonymous and nonsynonymous
substitutions \textcolor{red}{(Supplemental figure here)}. Together, these observations suggest that hard sweeps
are not a primary form of selection for either maize or teosinte.

@@

\subsection{Patterns of purifying and background selection can be explained by demography}
Purifying selection refers to the situation where deleterious
mutations arising in a population are continuously selected against. When this
form of selection is operating, it can serve to reduce genetic
variability at linked neutral sites, a phenomenon referred called
background selection \cite{charlesworth1993}. Purifying and background selection lead
to lower diversity within genes and other functional sites relative
to neutral regions \textcolor{red}{(ref here)}. We investigated purifying
selection in maize and teosinte by evaluating the average magnitude of reduced
diversity within genes and recovery away from genes in both taxa. When standardized by neutral
levels (which we defined as diversity at least \textcolor{red}{0.01} cM distal
from genes), a stronger reduction of diversity
and slower recovery was observed for teosinte than for maize, implying
that purifying selection has left a more pronounced signature in the
teosinte genome (Figure \ref{purify}). This conflicted with our \emph{a priori} hypothesis;
we expected that strong artificial selection since domestication would
have caused enhanced purifying selection for maize. We therefore
conducted the same analysis based on singleton diversity. As a class,
singleton alleles depict the most recent patterns of evolution, but
also have the lowest effect on pairwise diversity. Therefore, unlike
pairwise diversity, patterns of singleton
diversity reflect recent patterns of evolution. When
evaluating the data in this manner, an opposite relationship was
observed. Maize singleton diversity was lower than teosinte singleton
diversity near genes, and recovered more slowly
(Figure \ref{purify}), implying that in the
recent past maize has been more influenced by purifying selection than
teosinte. Together, these findings imply that demographic history has
a strong influence on the effect of purifying selection. Historically,
teosinte has had a larger population size than maize, and only
recently has maize population size overcome that of teosinte. Since
the efficacy of purifying selection scales with population size,
these results likely reflect changes in $N_e$ more than they reflect
underlying changes in selection pressure.

\begin{figure*}[b]
%\vspace*{.05in}
\centering
\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_manuscript.png}
\caption{Relative level of diversity versus distance to the nearest
  gene, in maize and teosinte. Two measures of diversity were
  investigated. \textbf{A} displays pairwise
  diversity, which is most influenced by intermediate frequency
  alleles and therefore depicts more ancient evolutionary patterns,
  and \textbf{B} depicts singleton diversity, influenced by rare
  alleles and thus depicting evolutionary patterns in the recent past.}
\label{purify}
\end{figure*}

@@

\bibliography{Reference}
\bibliographystyle{pnas}
%\nocite{*}
